---
title: "Readme for Excercise 3 of CFD Master Praktikum (Group G)"
author: "Yue Zhu, Yingqiang Gao and Nikhil Agarwal"
date: "June 4, 2018"
output: 
  pdf_document: 
    fig_caption: yes
    number_sections: yes
    toc: yes
  html_document: 
    fig_caption: yes
    number_sections: yes
    toc: yes
---
# Instructions about compilation, execution, and report generation

## Compilation and Execution

Please start a terminal and switch to where the `main.c` is. Build it first by typing `make` in the terminal. Then if everything is right, find the executable `sim` being created and run with the following command:

- `mpiexec -n <N> ./sim` for `MPICH` where `N` is the product of the number of MPI processes along x and y axis.
- `mpirun --hostfile my-hosts -np <N> ./sim` for `Open-MPI` where `N` is the product of the number of MPI processes along x and y axis.

*Example: To execute the binary with 2 process along x direction and 3 along y with MPICH, set these values in `cavity100.dat` and run as follows*

`mpiexec -n 6 ./sim`

## Report Generation

The markdown file (and generated html) loads image if the screenshot generated from the `Paraview` are stored at `.\..\output` location relative to README.rmd.

Use a markdown file viewer to view the report or it can be converted to html or pdf using suitable tools. `Github` automatically generates the view from markdown.

# Problem:

## Parameters

|            |            |              |                |
| ---------- | :--------: | :----------: | -------------: |
| imax = 300 | jmax = 300 | xlength = 10 | ylength = 10   |
| dt = 0.01  | t_end = 1  | tau = 0.5    | dt_value = 2.0 |
| eps = 0.01 | omg = 1.7  | alpha = 0.5  | itermax = 100  |
| GX = 0.0   | GY = 0.0   | Re = 10      |                |
| UI = 0.0   | VI = 0.0   | PI = 0.0     |                |
| iproc = 2  | jproc = 3  |              |                |

## Geometry

Driven cavity of square cross-section area. The domain is a container filled with a fluid with the container
lid (a band or a ribbon) moving at a constant velocity. No-slip conditions are imposed on all four boundaries, with the exception of the upper boundary, along which the velocity u in x-direction is not set to zero, but is equal to 1, in order to simulate the moving lid.

---

## Pressure

\begin{figure}[h!]
\includegraphics[width=0.75\textwidth]{./../output/output_i2j3_ex3parameters_pressure.png}
\caption{Pressure}
\end{figure}

**Observation**: The pressure is highest at the top right while lowest at top left, is consistent with the direction of fluid flow in the absence of thermal effects.

--- 

## Streamlines

\begin{figure}[h!]
\includegraphics[width=0.75\textwidth]{./../output/output_i2j3_ex3parameters_stream.png}
\caption{Streamlines}
\end{figure}

**Observation:** The streamlines are ploted for each subdomain independently. The path traced out by a massless particle as it moves with the flow, is matching at the interface of each subdomain. So, we can conclude that the streamlines are consistent with other subdomains.

\newpage
\pagebreak

## Velocity

\begin{figure}[h!]
\includegraphics[width=0.75\textwidth]{./../output/output_i2j3_ex3parameters_velocity.png}
\caption{Velocity}
\end{figure}


**Observation:** As expected velocity is maximum at the top. Due to low Reynold's number there is less turbulence in the system (**when  compared with Worksheet1**).

---

## Performance

- Speedup, `S(p) := T(1)/T(p)`
- Parallel efficiency `E(p) := T(1)=(p * T(p)) = S(p)/p`

On Intel Processor i7-8550U (*henceforth refered to as **new machine***) with

- Nominal Frequency = 1.80GHz
- Single Core Frequency = 4.0 GHz
- \# Core = 4
- \# Threads/core = 2
- L3 Cache = 8 MB
- Memory Types = DDR4-2400, LPDDR3-2133
- Bus Speed 4 GT/s OPI

| # Process (x) | # Process (y) | Total Process | time1 (s) | time2 (s) | time3 (s) | mean (s) | Std. Dev | Speedup | Efficiency |
| ------------- | ------------- | ------------- | --------- | --------- | --------- | -------- | -------- | ------- | ---------- |
| 1             | 1             | 1             | 164.37    | 165.75    | 165.27    | 165.00   | 0.70     | 1.00    | 1.00       |
| 2             | 1             | 2             | 88.52     | 88.86     | 91.43     | 90.00    | 1.59     | 1.83    | 0.92       |
| 1             | 3             | 3             | 35.19     | 36.28     | 36.16     | 36.00    | 0.59     | 4.58    | 1.53       |
| 1             | 4             | 4             | 31.51     | 30.12     | 26.85     | 30.00    | 2.39     | 5.50    | 1.38       |
| 1             | 6             | 6             | 35.37     | 36.66     | 35.91     | 36.00    | 0.65     | 4.58    | 0.76       |
| 2             | 3             | 6             | 39.48     | 39.18     | 40.58     | 40.00    | 0.74     | 4.13    | 0.69       |
| 2             | 4             | 8             | 35.51     | 45.52     | 39.96     | 40.00    | 5.02     | 4.13    | 0.52       |
| 2             | 2             | 4             | 53.78     | 52.98     | 55.17     | 54.00    | 1.11     | 3.06    | 0.76       |
| 3             | 4             | 12            | very slow |           |           |          |          |         |            |
| 4             | 5             | 20            | very slow |           |           |          |          |         |            |
| 6             | 6             | 36            | very slow |           |           |          |          |         |            |

---

On Intel Processor i7-4720HQ (_henceforth refered to as **old machine**_) with

- Nominal Frequency = 2.60GHz 
- Single Core Frequency = 3.6GHz
- \# Core = 4
- \# Threads/core = 2
- L3 Cache = 6 MB
- Memory Types = DDR3L 1333/1600
- Bus Speed 5 GT/s DMI2


| # Process (x) | # Process (y) | Total Process | time1 (s) | time2 (s) | time3 (s) | mean (s) | Std. Dev | Speedup | Efficiency |
| ------------- | ------------- | ------------- | --------- | --------- | --------- | -------- | -------- | ------- | ---------- |
| 1             | 1             | 1             | 115.13    | 112.08    | 114.09    | 113.77   | 1.55     | 1.00    | 1.00       |
| 2             | 1             | 2             | 60.56     | 58.85     | 61.00     | 60.13    | 1.14     | 1.89    | 0.95       |
| 1             | 3             | 3             | 42.97     | 43.86     | 44.00     | 43.61    | 0.56     | 2.61    | 0.87       |
| 1             | 4             | 4             | 36.93     | 33.60     | 36.60     | 35.71    | 1.83     | 3.19    | 0.80       |
| 1             | 5             | 5             | 39.74     | 42.55     | 44.18     | 42.16    | 2.24     | 2.70    | 0.54       |
| 1             | 6             | 6             | 40.89     | 42.86     | 45.99     | 43.25    | 2.57     | 2.63    | 0.44       |
| 2             | 4             | 8             | 45.42     | 50.26     | 52.96     | 49.55    | 3.82     | 2.30    | 0.29       |
| 2             | 2             | 4             | 36.34     | 34.51     | 36.50     | 35.78    | 1.11     | 3.18    | 0.79       |

---

\newpage
\pagebreak

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r speedup, echo=FALSE}
# Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

# Define 2 vectors
Speedup4  <- c(1.00,	1.89,	2.61,	3.19,	2.70,	2.63,	2.30)
Speedup8 <- c(1.00,	1.83,	4.58,	5.50,	4.58,	4.13,	4.13)

# Calculate range from 0 to max value of cars and trucks
g_range <- range(0, Speedup4, Speedup8)

# Graph using y axis that ranges from 0 to max 
# value in both y axis vector.  Turn off axes and 
# annotations (axis labels) so we can specify them ourself
plot(Speedup4, type="o", col="blue", ylim=g_range, axes=FALSE, ann=FALSE)

# Graph another y vector with red dashed line and square points
lines(Speedup8, type="o", col="red", ylim=g_range)

# Make x axis using labels
axis(1, at=1:7, lab=c(1,2,3,4,5,6,8))

# Make y axis with horizontal labels that display ticks at 
# every 4 marks. 4*0:g_range[2] is equivalent to c(0,4,8,12).
axis(2, las=1, at=1*0:g_range[2])
# axis(2, las=1, at=1*0:g_range[2])

# Create box around plot
box()

# Gridlines (https://stat.ethz.ch/R-manual/R-devel/library/graphics/html/grid.html)
grid(nx = NULL, ny = NULL, col = "lightgray", lty = "dotted", lwd = par("lwd"), equilogs = TRUE)

# Create a title with a red, bold/italic font
title(main="Speedup Comparison : Old Machine Vs New machine", col.main="Black", font.main=4)

# Label the x and y axes with dark green text
title(xlab="No. of Process", col.lab=rgb(0,0.5,0))
title(ylab="Speedup", col.lab=rgb(0,0.5,0))

# Create a legend at (1, g_range[2]) that is slightly smaller 
# (cex) and uses the same line colors and points used by 
# the actual plots 
legend(2, 1, c("Speedup with Old Machine","Speedup with New machine"), cex=0.8, 
       col=c("blue","red"), pch=21:22, lty=1:1)
```

---

```{r Efficiency, echo=FALSE}

# Define 2 vectors
Efficiency4  <- c(1, 0.95,	0.87,	0.8,	0.54,	0.44,	0.290)
Efficiency8 <- c(1.00,	0.92,	1.53,	1.38,	0.76,	0.69,	0.52)

# Calculate range from 0 to max value of cars and trucks
g_range <- range(0, Efficiency4, Efficiency8)

# Graph using y axis that ranges from 0 to max 
# value in both y axis vector.  Turn off axes and 
# annotations (axis labels) so we can specify them ourself
plot(Efficiency4, type="o", col="blue", ylim=g_range, axes=FALSE, ann=FALSE)

# Graph another y vector with red dashed line and square points
lines(Efficiency8, type="o", col="red", ylim=g_range)

# Make x axis using labels
axis(1, at=1:7, lab=c(1,2,3,4,5,6,8))

# Make y axis with horizontal labels that display ticks at 
# every 4 marks. 4*0:g_range[2] is equivalent to c(0,4,8,12).
axis(2, las=1, at=0.2*0:5*g_range[2])
# axis(2, las=1, at=1*0:g_range[2])

# Create box around plot
box()

# Gridlines (https://stat.ethz.ch/R-manual/R-devel/library/graphics/html/grid.html)
grid(nx = NULL, ny = NULL, col = "lightgray", lty = "dotted", lwd = par("lwd"), equilogs = TRUE)

# Create a title with a red, bold/italic font
title(main="Efficiency Comparison : Old Machine Vs New machine", col.main="Black", font.main=4)

# Label the x and y axes with dark green text
title(xlab="No. of Process", col.lab=rgb(0,0.5,0))
title(ylab="Efficiency", col.lab=rgb(0,0.5,0))

# Create a legend at (1, g_range[2]) that is slightly smaller 
# (cex) and uses the same line colors and points used by 
# the actual plots 
legend(2, 0.3, c("Efficiency with Old Machine","Efficiency with New machine"), cex=0.8, 
       col=c("blue","red"), pch=21:22, lty=1:1)
```

---


| **Observation**                                                                                                                                  | **Explanation**                                                                                                                     |
| ------------------------------------------------------------------------------------------------------------------------------------------------ | ----------------------------------------------------------------------------------------------------------------------------------- |
| Single Core performance of *Old machine* with lesser frequency, lesser cache, slow RAM but faster Bus is better than that of the *Newer machine* | There should me further CPU characteristics (beyond CPU, Cache, RAM, Bus) that should have caused this result                       |
|                                                                                                                                                  |                                                                                                                                     |
| When the total of Process becomes more the performance takes a severe hit!                                                                       | As the resources are limited, *oversubscription of process* leads to *severe performance penalty*.                                  |
|                                                                                                                                                  |                                                                                                                                     |
| Max. Speed up is achieved when iproc = 1 and jproc = 4                                                                                           | Communication overhead and resource contention affects speedup for higher number of MPI process.                                    |
|                                                                                                                                                  |                                                                                                                                     |
| Superlinearity in efficiency seen with the execution with the *newer machine*                                                                    | Probably due to more cache available in the machine (*Reference*: How to measure, present, and compare parallel performance - IEEE) |
|                                                                                                                                                  |                                                                                                                                     |
---

# Addendum

## Domain decomposition Visualized

See Figure *Domain Decomposition Visualized*

\begin{figure}[!h]
\includegraphics[width=0.75\textwidth]{./../output/output_i2j3_ex3parameters_proof_using_MPI.png}
\caption{Domain Decomposition Visualized}
\end{figure}

---

## Same Problem Visualized for Iproc = 2 and Jproc = 2

See Figure *Same Problem Visualized for Iproc = 2 and Jproc = 2*

\begin{figure}[!h]
\includegraphics[width=0.75\textwidth]{./../output/output_i2j2_ex3parameters_velocity.png}
\caption{Same Problem Visualized for Iproc = 2 and Jproc = 2}
\end{figure}

---

## Same Geometry but Parameters from Worksheet 1 Visualized for Iproc = 2 and Jproc = 2

See Figure *Same Geometry but Parameters from Worksheet 1 Visualized*

\begin{figure}[!h]
\includegraphics[width=0.75\textwidth]{./../output/output_i2j2_ex1parameters_velocity.png}
\caption{Same Geometry but Parameters from Worksheet 1 Visualized}
\end{figure}

---